---
title: "Group Assignment Time Series"
author: "MBD S2.2 Group 8"
date: "2/2/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```



# **THIS WILL BE CHANGED TO BE INTERACTIVE ONCE WE HAVE ALL THE TIME SERIES, IGNORE FOR NOW**
## TEST
</style>

<h3 class = 'menutop' id="Contents">
  [Contents](#Contents)
  [Introduction](#Introduction)

<div class="dropdown">
<a class="dropbtn">Data analysis</a>
<div class="dropdown-content">
  [Time series 1](#time_series_1)
  [Time series 2](#time_series_2)
  [Time series 3](#time_series_3)
  [Time series 4](#time_series_4)
</div>
</div>
  [Time series 5](#time_series_5)
  [Time series 6](#time_series_6)
  [Time series 7](#time_series_7)
</h3>
<body style = "font-family: Arial, Helvetica, sans-serif;">
<a center class="anchor" id="Introduction" style="background-color:Gainsboro; width:40%;"></a>
<center style="background-color:Gainsboro; width:40%;">Introduction</center>


> 'We are glad to present to you our first Group Assignment! We also wanted to practice R further, so we are delivering our findings in a markdown format.'

```{r, echo=F}
# Before starting, we have to install and activate some R libraries in order to use the necessary statistics and time series functions:
# basic statistics 
install.packages("fBasics")
# time series functions
install.packages("forecast")
# normality tests
install.packages("normtest")

# working with the excel tables
install.packages("xlsx")
install.packages("readxl")

library(fBasics)
library(forecast)
library(normtest)
library(tidyverse)
library(readxl)
#library(xlsx)


#clean environment
#rm(list=ls())
#File paths:
#Kiko file path
#data<-read.csv("C:/Users/usuario/Desktop/LPC/MASTERS/MBD_abril_2019/Sessions2&3sim.csv",header=TRUE,sep=";",dec=",")

#Oxana file paths
#file_path_original_data <- file.path("~","Downloads","Timeseries","Homework_1_DATA .csv")
#data<-read.csv(file_path_original_data, header=TRUE,sep=";",dec=",")
summary_table_ts_file <- file.path("~","Downloads","Timeseries","resultsfts.xlsx")
summary_table_ts_or <- read_excel(summary_table_ts_file, sheet = 1)
summary_table_ts_tr <- read_excel(summary_table_ts_file, sheet = 5)
tests_table <- read_excel(summary_table_ts_file, sheet = 2)

#Luna file paths
data<-read.csv(file_path_original_data, header=TRUE,sep=";",dec=",")

# Data itself
series1=data[,1] [1:300]
series2=data[,2] [1:300]
series3=data[,3] [1:300]
series4=data[,4] [1:300]
series5=data[,5] [1:2000]
series6=data[,6] 
series7=data[,7]
```

<div id= 'summary'>
# Summary of findings 

> 'We decided to also list the tests we used together with the original and transformed data'

## Original Data 

```{r date_variant, echo=FALSE}
print(summary_table_ts_or)
```

## Transformed Data

```{r date_variant, echo=FALSE}
print(summary_table_ts_tr)
```

## Tests

> 'Additionally we wanted to list all the tests we carried out in one place in order for it to be more understandable'

```{r date_variant, echo=FALSE}
print(tests_table)
```

</div>


<div id= 'time_series_1'>
## Time Series 1

```{r date_variant, echo=FALSE}

```
</div>


# HEY GUYS COPYPASTE FROM HERE TO '</div>' AND ADD YOUR SERIES INFO AND THE DESIRED GRAPHS I WILL MAKE IT LOOK NICE ON THRU
#AAAAAA
<div id= 'time_series_7'>
## Time Series 7
```{r date_variant, echo=FALSE}
y<-series7
# to adjust graphic size
par(mar=c(2,2,2,2)) 

# plot the series, its acf and pacf together
par(mfrow=c(3,1))
ts.plot(y)   
acf(y)
pacf(y)

# compute basic statistics
mean(y)
sd(y)
skewness(y)
kurtosis(y,method=c("moment"))  



```
</div>
